---
title: "Cleaning"
author: "Anuj Dahiya"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: pdf_document
---



The purpose of this document is to clean the `.csv` files stored in the *Step 2 - Reformat/Data csvs/* folder.

Before beginning, we need to import several packages that help run the code below.


```{r, message=FALSE, warning=FALSE}
library(knitr)
library(dplyr)
library(data.table)
library(lubridate)
```


# Access the data

In order to access the data we need to specify the folder we want to access the files from. We do this by defining a path to find them and set the path with `opts_knit$set(root.dir = path)`. In an R-Markdown document, it is required that you set the directory using this function. We can see the files in the directory with `head()`.

We also create a destination path with `dest` for later use so that we can export the files properly. `country_path` is a path to a dataset from which we will reference.

```{r}
path = "~/GitHub/FIDE/Chess Scripts/Step 2 - Reformat/Data csvs/"
country_path = "~/GitHub/FIDE/Chess Scripts/Step 3 - FIDE Country Codes/Country Data/FIDE_codes.csv"
dest =  "~/GitHub/FIDE/Chess Scripts/Step 4 - Cleaning/Cleaned csvs/"
opts_knit$set(root.dir = path)

temp = list.files(path = path)
full_path <- paste(path, temp, sep = "")

head(temp)
```



# Create functions to rename datasets

Below, I define a few functions that help us rename the datasets. 

`month` gets the first 3 letters of the temp elements, the month name
`num` converts each month into a number
`add_zero` converts each month number into a usable form when we create dates later on.

```{r}
month <- function(x){return(substr(x, 1, 3))}
num <- function(x) match(tolower(x), tolower(month.abb))
add_zero <- function(x){if (x <= 9){x = paste("0", x, sep = "")}; return(x)}
```


# Get the month number of all of the files in the dataset

Below, we rename create the variables names when they are assigned and imported.

```{r}
temp%>%
sapply(month)%>%
sapply(num)%>%
sapply(add_zero)%>%
paste("20", substr(temp, 4, 5), "-", ., "-", "01", sep = "")-> month_num

head(month_num)
```

We can see that the datasets correspond to dates on which the is recorded.


# Import all datasets

Below, we import and assign all of the datasets into memory. The only objects we need to hold onto is the data and several folder path references for later on. Therefore, we will remove everything unneeded with `rm(list=setdiff(ls())` below.

```{r, message=FALSE, warning=FALSE, eval = FALSE}
for(i in 1:length(full_path)) {
assign(month_num[i], fread(full_path[i], sep = "*", data.table = FALSE, 
                           strip.white = TRUE, blank.lines.skip = TRUE))
}

FIDE <- mget(ls(pattern = "[0-9][0-9]-[0-9][0-9]"))

rm(list=setdiff(ls(), c("FIDE", "path", "temp", "dest", "country_path")))
```


# Data prep

In order to get the data to have useful and common values, we need to rename dozens of columns and values. 

```{r, eval = FALSE}
vector_months <- c(month.abb, tolower(month.abb),toupper(month.abb))
string = ""
for (i in 1:length(vector_months)){
if (i == 1){string = vector_months[i]}
else if (i > 1) {string = paste(string, vector_months[i], sep = "|")}
}
string = paste(string, "RATING", sep = "|")


new = c("CM", "WCM", "WCM", "WGM", "WFM", "WFM", "GM", "IM", "FM", "WIM", "GM")
old = c("c", "wc", "WC", "wg", "WF", "wf", "g", "m", "f", "wm", "gm" )

dates = as.Date(names(FIDE))
months_vec <- months(dates)%>%toupper()%>%substr(., 1, 3) 
year_vec <- year(dates)%>%as.character()%>%substr(3,4)
files <- paste(months_vec, year_vec, ".csv", sep = "")

codes <- fread(country_path, 
               sep =  ",", header = TRUE)


country_codes <- c("BDI", "BHU", "BUR", "CAF", "CAM", "CGO", "CIV", "CMR", "COD", 
                   "CPV", "CUR", "DJI", "ERI", "FID", "FIE", "GAB", "GUM", "Ind", 
                   "IVC", "KOR", "KOS", "KSA", "LAO", "LBN", "LBR", "LCA", "LES", 
                   "MDV", "MTN", "NET", "NRU", "OMA", "PLW", "ROU", "SCG", "SGP", 
                   "SLE", "SOL", "SSD", "STP", "SWZ", "TLS", "TPE", "TTO") 

countries <- c("Berundi", "Bhutan", "Burkina Faso", "Central African Republic", "Cambodia", 
            "Republic of the Congo", "Cote d'Ivoire", "Cameroon", "Democratic Congo", 
            "Cape Verde", "Curaçao", "Djibouti", "Eritrea", "Finland", "FIE", "Gabon", 
            "Guam", "India", "Côte d'Ivoire", "South Korea", "Kosovo", "Saudi Arabia", 
            "Laos", "Lebanon", "Liberia", "Saint Lucia", "Lesotho", "Maldives", "Mauritania", 
            "NET", "Nauru", "Oman", "Palau", "Romania", "Serbia and Montenegro", "Singapore", 
            "Sierra Leonne", "Solomon Islands", "South Sudan", "Sao Tome and Principe", 
            "Swaziland", "East Timor (Timor-Leste)	", "Taiwan", "Trinidad and Tobago") 
```

# Ugly data cleaning

This is the ugly part of the document: a `for loop` that is really meaty. Essentially, we will iterate all of the FIDE datasets to adjust each data's columns and values. We need to do this because want common values to merge the data later. 

You are free to use `data.table`'s `rbindlist()` function to merge all of the datasets within `FIDE`, but I chose not to in this file.

```{r, results='hide', warning=FALSE, message=FALSE, eval=FALSE}
for(i in 1:length(FIDE)){
  colnames(FIDE[[i]])[grepl("Name|NAME|name", colnames(FIDE[[i]]))] <- "Name"
  colnames(FIDE[[i]])[grepl("NUMBER", colnames(FIDE[[i]]))] <- "ID_Number"
  colnames(FIDE[[i]])[grepl("Fed|FED|COUNTRY", colnames(FIDE[[i]]))] <- "Country"
  colnames(FIDE[[i]])[grepl("Gms|GAMES|GM|Game|GAME", colnames(FIDE[[i]]))] <- "Games"
  colnames(FIDE[[i]])[grepl("K", colnames(FIDE[[i]]))] <- "K_factor"
  colnames(FIDE[[i]])[grepl("FLAG|Flag|flag", colnames(FIDE[[i]]))] <- "Activity"
  colnames(FIDE[[i]])[colnames(FIDE[[i]]) %in% c("Wtit","wtit","WTIT", "WTit")] <- "Womens_Title"
  colnames(FIDE[[i]])[colnames(FIDE[[i]]) %in% c("TITLE","Title","title","Tit")] <- "Title"
  colnames(FIDE[[i]])[grepl(string, colnames(FIDE[[i]]))] <- "Rating"
  colnames(FIDE[[i]])[grepl("Born|Age|age|BIRTHDAY|B-day|Bday", colnames(FIDE[[i]]))] <- "Age_Birthday"
  colnames(FIDE[[i]])[grepl("SEX", colnames(FIDE[[i]]))] <- "Sex"
  colnames(FIDE[[i]])[grepl("FOA", colnames(FIDE[[i]]))] <- "FIDE_Online_Arena"
  colnames(FIDE[[i]])[grepl("OTit", colnames(FIDE[[i]]))] <- "Other_Titles"
  FIDE[[i]] <- FIDE[[i]] %>%
               mutate(Date = as.POSIXct(names(FIDE)[i], format="%Y-%m-%d"),
                      Date_numeric = year(Date)+yday(Date)/366,
                      Rating = as.numeric(Rating),
                      Title= c(new, Title)[match(Title, c(old, Title))],
                      Country = c(codes$Country, Country)[match(Country, c(codes$Code, Country))],
                      Country = c(countries, Country)[match(Country, c(country_codes, Country))])%>%
                      filter(!Country %in% c("Fed", "Col"))%>%
                      select(-one_of("V1"))
  fwrite(FIDE[[i]] , file = paste(dest, files[i], sep = ""), sep = "*")
}
```

# Example data after modifications

```{r, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
FIDE[[length(FIDE)]]%>%
head()%>%
select(Name, Country, Rating, Title, Date)%>%
kable()
```



```{r}
list.files(path = dest, pattern = "*.csv")%>%head()
```

